# Exploratory Data Analysis

This EDA comes as part of the model building process, where its focus are applicants based the following focus are termed as `good` or `bad` risks:

-   Job

-   Property

-   Credit History

-   employment

-   purpose

-   status

-   age

-   credit history

-   personal status

-   housing

## Importing libraries

```{r}
library(dplyr) # data manipulation & wrangling
library(moments) # statistical functions
library(ggplot2) # data visualization
```

### risk filtering

The goal is to filter the different risks and make assessment with them.

```{r}
good_risk <- credit_data_recode |> 
                filter(assessment == 'good')

dim(good_risk)
```

```{r}
# Bad risk
bad_risk <- credit_data_recode |> 
                filter(assessment == 'bad')

dim(bad_risk)
```

### Jobs Assessment

```{r}
jb_assess_bad <- table(bad_risk$job)
jb_assess_good <- table(good_risk$job)

print("Good Risks")
print(jb_assess_good)

print("Bad Risks")
print(jb_assess_bad)

```

```{r}
ggplot(credit_data_recode, aes(x = "", y = job, fill = job)) +
  geom_col() +
  coord_polar(theta = 'y') +
  labs(title = "Distribution of Loan Applicant Jobs") +
  theme_void()+
  theme(legend.position = 'bottom') 
```

```{r}
# Assuming your data is in a dataframe named 'credit_data'

# Create a summary table
summary_data <- credit_data %>%
  group_by(personal_status, assessment) %>%
  summarise(count = n())

# Create the pie chart
ggplot(summary_data, aes(x = "", y = count, fill = assessment)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  geom_text(aes(label = paste0(round(count/sum(count)*100), "%")), position = position_stack(vjust = 0.5)), group = assessment) +  # Group by assessment within each facet
  facet_wrap(~ personal_status) +
  labs(title = "Distribution of Risk Assessment by Marital Status")
```
